---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


# Loading and preprocessing the data

```{r load_data, echo=TRUE}
##Set Working Directory
setwd("D:/Document/RProgramming/RepData_PeerAssessment1")

##Unzip Data File
unzip("activity.zip")

##Read data from csv file##
activity <- read.csv("activity.csv", header=TRUE)

##Change dates to date variables##
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```

# What is mean total number of steps taken per day?

```{r question1, echo=TRUE}
## Get sums for days##
SumDays <- aggregate(activity$steps, by = list(activity$date), FUN = "sum")

##Assign Column Headers##
colnames(SumDays) <- c("Day", "Total_Steps")

## Draw Histogram##
library(ggplot2)
ggplot(SumDays, aes(x = Total_Steps)) + geom_histogram(binwidth = 800, fill="white", color="blue") + labs(title="Distribution of Total Steps Taken Every Day", x = "Steps per Day", y="Occurrences")

##Calculate Mean Number of Daily Steps##
MeanSteps <- mean(SumDays$Total_Steps, na.rm=TRUE)

##Calculate Media Number of Daily Steps##
MedianSteps <- median(SumDays$Total_Steps, na.rm=TRUE)
```
The Mean Daily Number of Steps is 'r MeanSteps'

The Median Daily Number of Steps is 'r MeadianSteps'

# What is the average daily activity pattern?

```{r question2, echo=TRUE}
##Drop Missing Values##
activity2 <- activity[!is.na(activity$steps), ]

##Calculate Interval Means##
Intervals <- aggregate(activity2$steps, by = list(activity2$interval), FUN = "mean")

##assign Column Headers##
colnames(Intervals) <- c("Time_Interval", "Mean_Steps")

##Draw Time Series Plot##
g <- ggplot(Intervals, aes(x = Time_Interval, y = Mean_Steps))
g + geom_line(binwidth = 800, fill="white", color="blue") + labs(title="Average Number of Steps Taken \n In 5 Minute Intervals Throughout the Day", x = "Time Interval", y = "Average Number of Steps")

##Find Interval with highest Number of Steps##
HighInterval <- Intervals[which.max(Intervals[, 2]), 1]
HighValue <- Intervals[which.max(Intervals[, 2]), 2]
```

The Maximum Number of Steps in a 5 minute interval is 'r HighValue' and this starts at 'r HighInterval' minutes into the day.

# Imputing missing values

```{r question3, echo=TRUE}

##Find the Number of Rows with Missing Data##
missing <- (nrow(activity)- sum(complete.cases(activity)))

##There are 'missing' rows that are missing data##

##Fill in Missing Values in New Dataframe with interval Averages##
activity3 <- merge(activity, Intervals, by.x = "interval", by.y = "Time_Interval", all.x = TRUE)
activity3$steps[is.na(activity3$steps)] <- activity3$Mean_Steps[is.na(activity3$steps)]
                 
##Get sums for days##
SumDays3 <- aggregate(activity3$steps, by = list(activity3$date), FUN = "sum")

##assign Column Headers##
colnames(SumDays3) <- c("Day", "Total_Steps")


##Plot the New Data##                 
library(ggplot2)
ggplot(SumDays3, aes(x = Total_Steps)) + geom_histogram(binwidth = 800, fill="white", color="blue") + labs(title="Distribution of Total Steps Taken Every Day", x = "Steps per Day", y="Occurrences")

##Calculate Mean Number of Daily Steps##
MeanSteps3 <- mean(SumDays3$Total_Steps)

##Calculate Media Number of Daily Steps##
MedianSteps3 <- median(SumDays3$Total_Steps)
```


The Mean Daily Number of Steps is 'r MeanSteps3'


The Median Daily Number of Steps is 'r MeadianSteps3'

# Are there differences in activity patterns between weekdays and weekends?

```{r question4, echo=TRUE}
##Find if day is weekdy or weekend##
activity3$Weekday <- ifelse(weekdays(activity3$date) != "Saturday" & weekdays(activity3$date) != "Sunday", "Weekday", "Weekend")

##Convert to Factor Data Type##
activity3$Weekday <- as.factor(activity3$Weekday)

##Find weekday means##
Intervals4 <- aggregate(activity3$steps, by = list(activity3$Weekday, activity3$interval), FUN = "mean")

##Assign Column Headers##
colnames(Intervals4) <- c("Weekday", "Interval", "Mean_Steps")

##Make the Plot##
ggplot(Intervals4, aes(Interval, Mean_Steps)) + 
    geom_line(binwidth = 800, fill="white", color="blue") + 
    facet_grid(Weekday ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```