---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


# Loading and preprocessing the data

Read the data file, unzip it, and change the date to date format.

```{r load_data, echo=TRUE}
setwd("D:/Document/RProgramming/RepData_PeerAssessment1")

unzip("activity.zip")

activity <- read.csv("activity.csv", header=TRUE)

activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```

# What is mean total number of steps taken per day?

Calculate total steps per day and plot a histogramof the distribution. Then post the mean and median.

```{r question1, echo=TRUE}

SumDays <- aggregate(activity$steps, by = list(activity$date), FUN = "sum")

colnames(SumDays) <- c("Day", "Total_Steps")

library(ggplot2)
ggplot(SumDays, aes(x = Total_Steps)) + geom_histogram(binwidth = 800, fill="white", color="blue") + labs(title="Distribution of Total Steps Taken Every Day", x = "Steps per Day", y="Occurrences")

MeanSteps <- mean(SumDays$Total_Steps, na.rm=TRUE)

MedianSteps <- median(SumDays$Total_Steps, na.rm=TRUE)
```

The Mean Daily Number of Steps is `r MeanSteps`

The Median Daily Number of Steps is `r MedianSteps`

# What is the average daily activity pattern?

Calculate the mean for each 5 minute period, then plot.  Find the period with the maximum average steps and post it.

```{r question2, echo=TRUE}

activity2 <- activity[!is.na(activity$steps), ]

Intervals <- aggregate(activity2$steps, by = list(activity2$interval), FUN = "mean")

colnames(Intervals) <- c("Time_Interval", "Mean_Steps")

g <- ggplot(Intervals, aes(x = Time_Interval, y = Mean_Steps))
g + geom_line(binwidth = 800, fill="white", color="blue") + labs(title="Average Number of Steps Taken \n In 5 Minute Intervals Throughout the Day", x = "Time Interval", y = "Average Number of Steps")

HighInterval <- Intervals[which.max(Intervals[, 2]), 1]
HighValue <- Intervals[which.max(Intervals[, 2]), 2]
```

The Maximum Number of Steps in a 5 minute interval is `r HighValue` and this starts at `r HighInterval` minutes into the day.

# Imputing missing values

Fill in the NA values with the mean for that interval, plot the histogram to see how it changes.  Post the mean and median.

```{r question3, echo=TRUE}

missing <- (nrow(activity)- sum(complete.cases(activity)))

activity3 <- merge(activity, Intervals, by.x = "interval", by.y = "Time_Interval", all.x = TRUE)
activity3$steps[is.na(activity3$steps)] <- activity3$Mean_Steps[is.na(activity3$steps)]
                 
SumDays3 <- aggregate(activity3$steps, by = list(activity3$date), FUN = "sum")

colnames(SumDays3) <- c("Day", "Total_Steps")

library(ggplot2)
ggplot(SumDays3, aes(x = Total_Steps)) + geom_histogram(binwidth = 800, fill="white", color="blue") + labs(title="Distribution of Total Steps Taken Every Day", x = "Steps per Day", y="Occurrences")

MeanSteps3 <- mean(SumDays3$Total_Steps)

MedianSteps3 <- median(SumDays3$Total_Steps)
```


The Mean Daily Number of Steps is `r MeanSteps3`


The Median Daily Number of Steps is `r MedianSteps3`

# Are there differences in activity patterns between weekdays and weekends?

Identify which days are weekdays and weekends, then post the data for each set of days to see if there is a difference in patterns.

```{r question4, echo=TRUE}

activity3$Weekday <- ifelse(weekdays(activity3$date) != "Saturday" & weekdays(activity3$date) != "Sunday", "Weekday", "Weekend")

activity3$Weekday <- as.factor(activity3$Weekday)

Intervals4 <- aggregate(activity3$steps, by = list(activity3$Weekday, activity3$interval), FUN = "mean")

colnames(Intervals4) <- c("Weekday", "Interval", "Mean_Steps")

ggplot(Intervals4, aes(Interval, Mean_Steps)) + 
    geom_line(binwidth = 800, fill="white", color="blue") + 
    facet_grid(Weekday ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```